---
title: "The Office (US) - how negative was the show?"
description: "Performing a bit of EDA on The Office transcript and IMDB ratings, along with sentiment analysis of the characters"
output: 
  md_document:
    fig_width: 13
    fig_height: 8
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message=FALSE,
                      warning=FALSE,
                      cache = TRUE
)
```


When I first saw this dataset come up on [TidyTuesday](https://github.com/rfordatascience/tidytuesday) the main thing I wanted to do was recreate the graph made by [pudding.cool](https://pudding.cool/2017/08/the-office/) where they plot character sentiment by season, prompting me to use Plotly for the first time. What I like about the chart is how it cleanly shows the character's sentiment level throughout all seasons (where they had a threshold of >30 lines) and also portrays how negative the show is overall. That said, whilst working on this I became slightly obsessed and started doing all kinds of different EDA.


```{r initialchunk, include=FALSE}
library(schrute)
library(forcats)
library(ggrepel)
library(tidytext)
library(ggthemes)
library(textdata)
library(ggridges)
library(reshape2)
library(wordcloud)
library(plotly)
library(tidyverse)



remove_regex <- "[:punct:]|[:digit:]|parts |part |the |and"

ratings_raw <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-17/office_ratings.csv') 

office_ratings <- ratings_raw # %>%
  # transmute(
  #   episode_name = str_to_lower(title),
  #   episode_name = str_remove_all(episode_name, remove_regex),
  #   episode_name = str_trim(episode_name),
  #   imdb_rating
  # )

  
office_transcript <- as_tibble(schrute::theoffice) %>%
  mutate(
    season = as.integer(season),
         episode = as.integer(episode)
    )



main_characters <- c("Angela", "Kelly", "Toby",
                     "Phyllis", "Oscar", "Dwight",
                     "Darryl", "Jan", "Ryan",
                     "Pam", "Kevin","Michael",
                      "Erin", "Jim", "Andy" , 
                     "Other")

transcript_words <- office_transcript %>%
  select(-text_w_direction, -total_votes, -air_date) %>%
  unnest_tokens(word, text) %>%
  dplyr::anti_join(stop_words, by = "word")
```


**To start my attempt**, we *inner join* the sentiments tibble we got through tidytext. In this case, I use the Bing Liu lexicon to classify it as just negative or positive. I then use *spread* to give negative and positive their own column so I can calculate overall sentiment by subtracting one from the other.
```{r char_sentiment_season}
sentiments = get_sentiments("bing")

character_sentiment_season <- transcript_words %>%
  inner_join(sentiments, by = "word") %>% 
  count(season, character, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  filter(character %in% main_characters &
           negative+positive > 50) %>%
  mutate(sentimentc = positive - negative)

character_sentiment_season
```

The below chunk took a bit of fiddling to get right, however, the main aim was to recreate the plot in a static way before calling *ggplotly* on it. *avg_max_sentiment* here was created to:

1. avg_sentiment used for ordering and to colour a character's average sentiment 

2. max_sentiment used for positioning of *geom_text*
```{r char_sentiment_season_graph}
avg_max_sentiment <- character_sentiment_season %>%
  group_by(character) %>%
  summarise(avg_sentiment = mean(sentimentc),
            max_sentiment = max(sentimentc),
            plotly_pos = max(sentimentc) + 3)

character_sentiment_season_adj <- character_sentiment_season %>%
  inner_join(avg_max_sentiment, by="character")


# static plot completed
p <- character_sentiment_season_adj %>%
  mutate(sent_dummy = ifelse(sentimentc < 0, "More Negative", "More Positive")) %>%
  mutate(character = reorder(character, avg_sentiment)) %>%
  ggplot(aes(character, sentimentc, hoverinfo = season)) +
  geom_point(colour = "#666666", size = 4, fill = "#f2f2f2", shape = 21) +
  coord_flip() +
  geom_hline(yintercept = 0, alpha = 0.2) +
  geom_point(aes(character, avg_sentiment, color=avg_sentiment), shape = 21, colour = "#262626", fill = "#ff9933", size = 4) + 
  geom_text(aes(label = character), 
                  size = 3, 
                  data = subset(character_sentiment_season_adj, sentimentc == max_sentiment),
            hjust = -0.5,
            vjust = 0.3) +
  labs(title = "How negative are The Office (US) characters?",
       subtitle = "Sentiment of each character based on Bing sentiment scores") +
  annotate(
    geom = "text",
    x = "Andy",
    y = -70,
    label = "More Negative",
    size = 3,
    colour = "#666666",
    alpha = .4
  ) +
  annotate(
    geom = "text",
    x = "Andy",
    y = 70,
    label = "More Positive",
    size = 3,
    colour = "#2b991f",
    alpha = .4
  ) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        panel.background = element_rect(fill = "white"),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(size = 6)
        )

p
```

Overall this comes pretty close to the chart created by pudding.cool - without the interactivity, of course. The main challenges with creating it were around getting the grid to disappear (and stay gone), especially after going through knitr. Additionally, creating the annotations for 'More Negative' and 'More Positive' could be done better but this is what I landed on.  In terms of making it interactive, I tried creating a plotly as per the below:
```{r interactive_char_sentiment_season_graph, dependson = "interactive_char_sentiment_season"}
ggplotly(p,
         tooltip = "season",
         width = 1000,
         height = 800) %>%
  style(textposition="right")
```

There are plenty of issues here that need solving, however, for the sake of this exercise I've left it where it was (especially as ggplotly currently does not support hjust/vjust).  

With that out of the way, I wanted to look at some other things. As someone who is not a fan of the series in any way, I was interested in finding out what seasons were successful and which episodes were fan favourites. 

### EDA
```{r echo=FALSE, top20_episodes_office}
office_ratings %>%
  arrange(desc(imdb_rating)) %>%
  mutate(title = paste0(season, ".", episode, " ", title),
         title = fct_reorder(title, imdb_rating)) %>%
  head(20) %>%
  ggplot(aes(title, imdb_rating, color = factor(season), size = total_votes)) +
  geom_point() +
  coord_flip() +
  labs(x = "IMDB rating", 
       y = "Episode",
       title = "Top 20 episodes of The Office",
       subtitle = "Based on IMDB ratings, colour represents season, size is # of votes") +
  theme(legend.position = "none") 

```

To me it's quite interesting that there are high rated episodes in a variety of seasons. The only two seasons that don't seem to feature in the top20 are season 1 and 8 (which judging by the reviews I quickly read on metacritic, weren't as well received as a whole anyway). Season 7 is at the top with 4 top rated episodes. Now that we've done the best, it would be interesting to see if Season 1 and 8 appear in the worst...

```{r echo=FALSE, bottom20_episodes_office}
office_ratings %>%
  arrange(imdb_rating) %>%
  mutate(title = paste0(season, ".", episode, " ", title),
         title = fct_reorder(title, imdb_rating)) %>%
  head(20) %>%
  ggplot(aes(title, imdb_rating, color = factor(season), size = total_votes)) +
  geom_point() +
  coord_flip() +
  labs(x = "IMDB rating", 
       y = "Episode",
       title = "Worst 20 episodes of The Office",
       subtitle = "Based on IMDB ratings, colour represents season, size is # of votes") +
  theme(legend.position = "none")
```

Season 8 and 9 both have a high share on this list. What stands out to me as well is that the Pilot is one of the lower rated episodes but people still kept watching. Even though looking at episodes by themselves is interesting, they definitely don't tell the whole story. Let's create a line graph so we can get an idea of how each season performed and what their best / worst episodes were.

```{r echo=FALSE, office_rating_by_episode_annotated}
office_ratings %>%
  mutate(title = fct_inorder(title),
         episode_number = row_number()) %>%
  ggplot(aes(episode_number, imdb_rating)) +
  geom_line() +
  geom_point(aes(color = factor(season), size = total_votes)) +
  geom_text(aes(label = title), check_overlap = TRUE, hjust = 1) +
  expand_limits(x = -10) +
  theme(panel.grid.major.x = element_blank(),
        legend.position = "none") +
  labs(x = "Episode number",
       y = "IMDB rating",
       title = "Popularity of The Office episodes over time",
       subtitle = "Colour represents season, size represents number of votes")
```

There's a lot going on here but it does immediately give an idea of how the series did throughout its lifetime. My immediate questions would be why did people hate *The Banker* and *Get the Girl* so much? Looking it up, turns out that people hated *The Banker* because it was a clip episode and *Get the Girl* had Catherine Tate play a character that had very mixed reviews. As someone who isn't a fan of the show, I also didn't realise Steve Carell left but the *Goodbye, Michael* episode immediately stands out (especially with a rating of 9.7). Of course, you can also see the perceived quality of the show drops with Season 8 looking like a complete failure reviews-wise, and Season 9 only pulling it back with the final few episodes.

As the line chart is pretty convoluted, I thought I'd try doing a circular bar chart for the first time following the guidelines at [r-graph-gallery](https://www.r-graph-gallery.com/297-circular-barplot-with-groups.html). In the below chunk, I:

1. create a dataframe with a filter on the review scores - this will create a column where I can easily check the condition

2. use *case_when* to pull out title if condition is true, leave blank if not

3. create a few variables, all of the ones here come straight from the article above

4. had to specify a font as default was causing issues

5. create graph

```{r office_ratings_circular_bar}
office_bar_chart_data <- data.frame(1:nrow(office_ratings), office_ratings, office_ratings$imdb_rating >= 9 | office_ratings$imdb_rating <= 7.4)
names(office_bar_chart_data)[1] <- "id"
names(office_bar_chart_data)[8] <- "condition"


office_bar_chart_data <- office_bar_chart_data %>%
  mutate(rated_title = case_when(condition == TRUE ~ title,
                             condition == FALSE ~ ""))


# label
label_data <- office_bar_chart_data
number_of_bars <- nrow(label_data)
angle <- 90 - 360 * (label_data$id - 0.5) / number_of_bars
label_data$hjust <- ifelse(angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle + 180, angle)

windowsFonts(Arial=windowsFont("TT Arial"))
font <- "Arial"

p <- ggplot(office_bar_chart_data, aes(x = as.factor(id),
                                y = imdb_rating,
                                fill = as.factor(season))) +
  geom_bar(stat="identity") +
  coord_polar(start = 0) +
  ylim(-10, 15) +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "The Office (US) episodes with rating",
       subtitle = "Colour represents Season, titles shown for episodes with score equal or greater than 9 OR scores below 7.5") +
  theme_minimal() +
  theme(text = element_text(family = font),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank(),
        plot.margin = unit(c(0, -5, 0, -5), "cm"),
        legend.position = "none",
        plot.title = element_text(face = "bold",
                                  colour = "black", size = 16, hjust = 0.5),
        plot.subtitle = element_text(face = "italic",
                                     colour = "black", size = 11, hjust = 0.5)
  )


p + geom_text(data=label_data, 
              aes(x=id, 
                  y=imdb_rating, 
                  label=rated_title, 
                  hjust=hjust), 
              color="black", 
              family=font,
              fontface="bold",
              alpha=0.6, 
              size=2.5, 
              angle = label_data$angle, 
              inherit.aes = FALSE ) 
```

The graph itself would perhaps not really work in the real world as it's quite hard to see general trend. That said, it's quite good tos ee the top/bottom rated episodes especially with the titles appended to it. The main struggle with circular bar plot, for me, is getting the angle right but the article linked was quite helpful. We did episode-specific so I quickly wanted to a *facet* of seasons too, to show the drops.

```{r echo=FALSE, office_ratings_through_seasons}
average_rating <- office_ratings %>%
summarise(avg_rating = mean(imdb_rating))

office_ratings %>%
ggplot(aes(episode, imdb_rating)) +
geom_line() +
geom_hline(yintercept = 8.26) +
facet_wrap(~ season) +
  labs(title = "The Office Ratings, all seasons",
       subtitle = "horizontal line represents the average rating, 8.26") +
  theme(
        plot.title = element_text(face = "bold",
                                  colour = "black", size = 16, hjust = 0.5),
        plot.subtitle = element_text(face = "italic",
                                     colour = "black", size = 11, hjust = 0.5)
  )
```

Important to note that Season 1 had fewer episodes which is why the line is much shorter. Using the average, we can see that season 1 barely lived up to what was yet to come and the show seems to peak between season 2 - 5. It now also becomes clearer what effect Steve Carrell leaving had on the series, with Season 8 not having a single episode with an average rating or higher, and Season 9 only having 4. Lastly for this section I wanted to have a quick look at what character has the highest average rating.

```{r include=FALSE, summariseratings}
ratings_summarised <- office_ratings %>%
  group_by(title) %>%
  mutate(imdb_rating = mean(imdb_rating))

character_lines_rating <- office_transcript %>%
  count(character, episode_name) %>%
  group_by(character) %>%
  filter(sum(n) >= 100) %>%
  left_join(ratings_summarised, by = c("episode_name" = "title")) %>%
  drop_na()     # this will drop the part 1 / part 2
```


```{r char_lines_rating, dependson="summariseratings"}
character_lines_rating %>%
  summarise(
    avg_rating = mean(imdb_rating),
    nb_episodes = n()
  ) %>%
  arrange(desc(avg_rating)) %>%
  ggplot(aes(fct_reorder(character, avg_rating), avg_rating, size = nb_episodes)) +
  geom_point(show.legend = FALSE) +
  coord_flip() +
  labs(x = "Character name",
       y = "Average rating",
       title = "Which character in The Office has the highest average episode rating?",
       subtitle = "Size represents number of episodes featured")
```


Turns out that Charles, who is only in 7 episodes, has the highest average rating. After having a look at who he is, it's a character played by Idris Elba and was received very well by the audiences. Michael also has a high average episode rating and it probably helps that he didn't appear in the later 2 seasons (except for last episode).

### Text Analysis

So, what are the most commonly used words in The Office?

```{r office_common_words}
transcript_words %>%
  count(word, sort=TRUE) %>%
  head(20) %>%
  ggplot(aes(fct_reorder(word, n), n)) +
  geom_col() +
  coord_flip()
```

So apparently characters really like saying each other's name which isn't that surprising in an office environment. To round it off, a quick wordcloud showing the most used words in the office. I split them by positive and negative because as the very first chart showed, there's definitely more negativity than positivity in the show. That said, the most commonly used word is, apparently, "love".

```{r office_word_cloud}
office_sentiment_count <- transcript_words %>%
  inner_join(sentiments %>%
               filter(sentiment == "positive"|
                      sentiment == "negative")) %>%
  count(word, sentiment, sort = TRUE)

office_sentiment_count %>%
  acast(word~sentiment, value.var='n', fill = 0) %>%
  comparison.cloud(colors=c("#F8766D", "#00BFC4"), 
                   max.words=200)
```